\bf Partial Least Squares Regression with Multiple Functional and Scalar Predictors

### 2025-07-11 create-FSHybridPLS.Rmd
code:get_xi_direct removed

```{r}
get_xi_direct <- function(W, y) {
  K <- W$n_functional
  C_J <- get_CJ(W)
  A <- cbind(C_J, W$Z)                      # shape: n × d
  b <- t(y) %*% A / W$n_sample   # equivalent to V_star

  J_star <- get_gram_matrix_block(W)
  alpha <- solve(J_star, t(b))                # unconstrained maximizer
  norm_squared <- as.numeric( t(alpha) %*% J_star %*% alpha )
  xi_hat <- alpha / sqrt(norm_squared)  # normalize

  return(xi_hat)
}
```

@@

# --- Step 1: Simulate Data ---
set.seed(777)

>>>>

@@

# --- Step 1: Simulate Data ---
set.seed(2222)

### 2025-07-11 manuscript/main.tex
code:get_xi_direct removed

Let $\hat{\mathbf{d}}_j := (\hat{d}_{j1}, \ldots, \hat{d}_{jM})^\top \in \mathbb{R}^M$ and $\mathbf{y} := (y_1, \ldots, y_n)^\top$. 
Let $B \in \mathbb{R}^{M \times M}$ denote the Gram matrix of the basis functions, with entries
\[
B_{m, m'} := \langle b_m, b_{m'} \rangle, \quad m, m' = 1, \ldots, M.
\]
We compute $\hat{\mathbf{d}}_1, \ldots, \hat{\mathbf{d}}_K$ and $\hat{\boldsymbol{\zeta}}$ as 
\begin{equation*}
	\hat{\mathbf{d}}_j  = \frac{1}{\sqrt{\sum_{j=1}^K 
			\mathbf{y}^\top
			\Theta_j
			B \Theta_j^\top \mathbf{y} 
			+ 
			\mathbf{y}^\top Z Z^\top \mathbf{y}}}
	\Theta_j^\top \mathbf{y},
	\quad
	%%%%%%%%%%%%%%%%
	\hat{\boldsymbol{\zeta}}  = \frac{1}{\sqrt{\sum_{j=1}^K 
			\mathbf{y}^\top
			\Theta_j
			B \Theta_j^\top \mathbf{y} 
			+ 
			\mathbf{y}^\top Z Z^\top \mathbf{y}}}
	Z^\top \mathbf{y}.
\end{equation*}
The derivation of this formula is provided in Appendix~\ref{section:proof:eigenproblem_derivation}.
Although the unnormalized coefficients are computed separately for the functional and scalar part of the PLS component direction, the normalization step couples them, allowing the components to influence each other. This enables the procedure to account for the correlation between functional and scalar components.

@@

\subsection{Proof of Lemma \ref{lemma:valid_inner_product_tuple}}\label{section:proof:lemma:valid_inner_product_tuple}
\begin{proof}
	We verify that the function
	\[
	\langle u, v \rangle_{B} := \sum_{j=1}^K \mathbf{d}_j^\top B \mathbf{g}_j + \boldsymbol{\zeta}^\top \boldsymbol{\omega},
	\]
	defines a valid inner product on $ (\mathbb{R}^M)^K \times \mathbb{R}^p$, where $u = (\mathbf{d}_1, \ldots, \mathbf{d}_K, \boldsymbol{\zeta})$ and $v = (\mathbf{g}_1, \ldots, \mathbf{g}_K, \boldsymbol{\omega})$ are elements of $(\mathbb{R}^M)^K \times \mathbb{R}^p$.
	\paragraph{Symmetry}: Since $B$ is symmetric and the Euclidean inner product is symmetric,
	\[
	\langle u, v \rangle_{B} = \sum_{j=1}^K \mathbf{d}_j^\top B \mathbf{g}_j + \boldsymbol{\zeta}^\top \boldsymbol{\omega} = \sum_{j=1}^K \mathbf{g}_j^\top B \mathbf{d}_j + \boldsymbol{\omega}^\top \boldsymbol{\zeta} = \langle v, u \rangle_{B}.
	\]
	\paragraph{Linearity in the first argument}: Let $u^{(1)} = (\mathbf{d}_1^{(1)}, \ldots, \mathbf{d}_K^{(1)}, \boldsymbol{\zeta}^{(1)})$ 
	and
	$v = (\mathbf{g}_1, \ldots, \mathbf{g}_K, \boldsymbol{\omega}) $
	be the elements of $(\mathbb{R}^M)^K \times \mathbb{R}^p$.
	Let $a, b \in \mathbb{R}$. Then we have
	\begin{align*}
		\langle a u^{(1)} + b u^{(2)}, v \rangle_{B}
		&= \sum_{j=1}^K (a \mathbf{d}_j^{(1)} + b \mathbf{d}_j^{(2)})^\top B \mathbf{g}_j + (a \boldsymbol{\zeta}^{(1)} + b \boldsymbol{\zeta}^{(2)})^\top \boldsymbol{\omega} \\
		&= 
		a  \sum_{j=1}^K (\mathbf{d}_j^{(1)})^\top B \mathbf{g}_j 
		+
		b \sum_{j=1}^K  (\mathbf{d}_j^{(2)})^\top B \mathbf{g}_j
		+
		a ( \boldsymbol{\zeta}^{(1)})^\top \boldsymbol{\omega}
		+
		b ( \boldsymbol{\zeta}^{(2)})^\top \boldsymbol{\omega}
		\\
		&= 
		a  
		\biggl(
		\sum_{j=1}^K (\mathbf{d}_j^{(1)})^\top B \mathbf{g}_j 
		+
		( \boldsymbol{\zeta}^{(1)})^\top \boldsymbol{\omega}
		\biggr)
		+
		b 
		\biggl( 
		\sum_{j=1}^K  (\mathbf{d}_j^{(2)})^\top B \mathbf{g}_j
		+
		( \boldsymbol{\zeta}^{(2)})^\top \boldsymbol{\omega}
		\biggr)
		\\
		&= a \langle u^{(1)}, v \rangle_{B} + b \langle u^{(2)}, v \rangle_{B}.
	\end{align*}

	\paragraph{Positive definiteness.}: we have
	$
	\langle u, u \rangle_{B} = \sum_{j=1}^K \mathbf{d}_j^\top B \mathbf{d}_j + \boldsymbol{\zeta}^\top \boldsymbol{\zeta}
	$
	Since $B$ is positive semi-definite, each term $\mathbf{d}_j^\top B \mathbf{d}_j \geq 0$, and $\boldsymbol{\zeta}^\top \boldsymbol{\zeta} \geq 0$, with equality if and only if $\mathbf{d}_j = 0$ for all $j$ and $\boldsymbol{\zeta} = 0$. This completes the proof of Lemma \ref{lemma:valid_inner_product_tuple}.
\end{proof}

>>>>

	Let $\hat{\mathbf{d}}_j := (\hat{d}_{j1}, \ldots, \hat{d}_{jM})^\top \in \mathbb{R}^M$ and $\mathbf{y} := (y_1, \ldots, y_n)^\top$. 
	Let $B \in \mathbb{R}^{M \times M}$ denote the Gram matrix of the basis functions, with entries
	\[
	B_{m, m'} := \langle b_m, b_{m'} \rangle, \quad m, m' = 1, \ldots, M.
	\]
	We compute $\hat{\mathbf{d}}_1, \ldots, \hat{\mathbf{d}}_K$ and $\hat{\boldsymbol{\zeta}}$ as 
	\begin{equation*}
		\hat{\mathbf{d}}_j  = \frac{1}{\sqrt{\sum_{j=1}^K 
				\mathbf{y}^\top
				\Theta_j
				B \Theta_j^\top \mathbf{y} 
				+ 
				\mathbf{y}^\top Z Z^\top \mathbf{y}}}
		\Theta_j^\top \mathbf{y},
		\quad
		%%%%%%%%%%%%%%%%
		\hat{\boldsymbol{\zeta}}  = \frac{1}{\sqrt{\sum_{j=1}^K 
				\mathbf{y}^\top
				\Theta_j
				B \Theta_j^\top \mathbf{y} 
				+ 
				\mathbf{y}^\top Z Z^\top \mathbf{y}}}
		Z^\top \mathbf{y}.
	\end{equation*}
These expressions represent matrix–vector multiplications normalized by a common factor. 
	The derivation of this formula is provided in Appendix~\ref{section:proof:Proposition:linear_system_unregularized}.
	Although the unnormalized coefficients are computed separately for the functional and scalar part of the PLS component direction, the normalization step couples them, allowing the components to influence each other. This enables the procedure to account for the correlation between functional and scalar components.

@@
